---
title: "PM2.5 Tutorial"
author: "Yuming Liu"
date: "October 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

## Load the Packages

We can use the "Packages" set-up on the right hand side to install the packages, or we could just run the code # install.packages('name', dependencies=TRUE, repos='http://cran.rstudio.com/')# from the Console to install new packages.

Spatial analysis in R requires multiple libraries. Package installation is done with the following syntax: install.packages(“sp”). Some of these take additional time for installation, depending on your system. The following list is comprehensive for this tutorial, as well as much spatial analysis.

```{r packages,include=FALSE}
library("sp") #spatial data wrangling & analysis
library("rgdal")
library("rgeos") #spatial data wrangling & analysis
library(raster) #spatial data wrangling & analysis
library(adehabitatHR) # Spatial data wrangling
library(tmap)  #modern data visualizations
library(data.table)  
library(gstat) #spatial data wrangling & analysis
library(xts) #data wrangling & analysis
```

## Loading the Shapefiles

For ploting the PM2.5, we need to read the community area shapefiles and the shapefiles of the epa sensors.

```{r shapefiles1}
Chicago <- readOGR(".", "CommAreas")
plot(Chicago)
```
```{r shapefiles2}
EPA.Points <- readOGR(".", "PM2.5YearlyShapefile_CLEAN")
str(EPA.Points@data)
EPA.Points@data
plot(EPA.Points)
```

## Filtering the Sensors

We need to take out the sensor points that are located too far from Chicago.

```{r filter}
test <- EPA.Points
test<- test[-c(24,25), ]
str(test@data)
plot(test)
EPA.Points <- test
```


## Generate the IDW

To set up the grid, we define the sample grid based on the extent of the "EPA.Points" file.

```{r}
grid <-spsample(EPA.Points, type = 'regular', n = 10000)
```

Then we run the IDW function from the gstat package for the variable of interest in EPA.Points.

```{r}
idw <- idw(EPA.Points$X2017MPC ~ 1, EPA.Points, newdata = grid)

idw.output = as.data.frame(idw)
names(idw.output)[1:3] <- c("long", "lat", "prediction")
```

Then we create the spatial points data frame.

```{r}
spg <- idw.output
coordinates(spg) <- ~ long + lat
```

The last step is to coerce the grid to Spatial Pixels Data Frame and to raster.

```{r}
gridded(spg) <- TRUE

raster_idw <- raster(spg)
```

## Set the Projection

We use the following code to set the standard of the projection.

```{r}
projection(raster_idw) <- CRS("+init=epsg:4326")
```

## Plot the IDW Results

Firstly, we mask raster by Chicago polygon file.

```{r}
masked_idw <- mask(raster_idw, Chicago)
```

Then we plot the masked raster.

```{r}
tm_shape(masked_idw) + tm_raster("prediction", style = "quantile", n = 100, legend.show = FALSE)
```

The previous raster file only shows the IDW result. Then We plot the masked raster with point emissions.

```{r}
tm_shape(masked_idw) + tm_raster("prediction", style = "quantile", n = 100, legend.show = FALSE) +
  tm_shape(EPA.Points) + tm_bubbles(size = "X2017MPC", col = "X2017MPC", palette = "Blues", 
                                    scale = 0.5, style = "quantile", legend.size.show = FALSE, title.col = "2017 PM2.5 Mean Concentration")

tmap_mode("view")

tm_shape(masked_idw) + 
  tm_raster("prediction", style = "quantile", 
            n = 100, alpha = 0.4, legend.show = FALSE) + 
  tm_shape(EPA.Points) + 
  tm_bubbles(col = "X2017MPC", 
             palette = "Blues", 
             scale = .5, 
             style = "quantile", 
             title.col = "2017 Mean PM2.5") 

tmap_mode("plot")
```

We plot the masked raster with Community Area overlay added.

```{r}
tm_shape(masked_idw, bbox=Chicago) + 
  tm_raster("prediction", style = "quantile", pal = "Blues", n = 5, legend.show = TRUE, title = "PM2.5 prediction") + 
  tm_shape(Chicago) + tm_borders(alpha=.5,) + 
  tm_layout(frame = F, legend.position = c("left","bottom"))

```

## Combine the shapefiles and the PM2.5 values

Extracting the Pixel Values from the Raster for chicago Community Areas and removing NAs, we have the cleaned masked IDW results.

```{r}
chi.ave <- raster::extract(masked_idw, Chicago)

chi.ave <- lapply(chi.ave, na.omit)
```

Then we calculate the average value by Community Areas.

```{r}
ca.ave <- lapply(chi.ave, FUN=mean)
ca.ave  <- unlist(ca.ave )
```

We attach the Mean Values to Community Area information by grabing Community Area names/IDs from Chicago shapefiles. 

```{r}
chi.info <- cbind(as.integer(as.character(Chicago@data$area_numbe)))
head(chi.info)
```

Then we attach EPA Average numbers to the data frame.

```{r}
chica.final <- cbind(chi.info, ca.ave)
colnames(chica.final) <- c("area_numbe", "PM2.5")
as.data.frame(chica.final)

Chicago<- merge(Chicago, chica.final, by = "area_numbe")
head(Chicago@data)
```

Now we plot the masked raster with Community Area overlay.

```{r}
tm_shape(Chicago) + 
  tm_fill("PM2.5", style = "quantile", n = 5, pal = "Blues",legend.show = TRUE, title = "PM2.5 by CA") + 
  tm_shape(Chicago) + tm_borders(alpha=.5,) + 
  tm_layout(frame = F, legend.position = c("left","bottom"))
```

## Save the Results

Finally, we write the data to a CSV file. 

```{r}
write.csv(chica.final, "PM5_aggregated.csv")
```

