---
title: "PM2.5 Tutorial"
author: "Yuming Liu"
date: "October 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

## Load the Packages

We can use the "Packages" set-up on the right hand side to install the packages, or we could just run the code `install.packages('name', dependencies=TRUE, repos='http://cran.rstudio.com/')` from the Console to install new packages.

Spatial analysis in R requires multiple libraries. Package installation is done with the following syntax: `install.packages(“sp”)`. Some of these take additional time for installation, depending on your system. The following list is comprehensive for this tutorial, as well as much spatial analysis.

```{r packages,include=FALSE}
library("sp") #spatial data wrangling & analysis
library("rgdal")
library("rgeos") #spatial data wrangling & analysis
library(raster) #spatial data wrangling & analysis
library(adehabitatHR) # Spatial data wrangling
library(tmap)  #modern data visualizations
library(data.table)  
library(gstat) #spatial data wrangling & analysis
library(xts) #data wrangling & analysis
```

## Loading the Shapefiles

For ploting the PM2.5, we need to read the community area shapefiles and the shapefiles of the epa sensors. We firstly plot the shapefiles for CHicago Community Areas and then plot the EPA sensors. 

```{r shapefiles1}
Chicago <- readOGR(".", "CommAreas")
plot(Chicago)
```
```{r shapefiles2}
EPA.Points <- readOGR(".", "PM2.5YearlyShapefile_CLEAN")
str(EPA.Points@data)
EPA.Points@data
plot(EPA.Points)
```

## Filtering the Sensors

We need to take out the sensor points that are located too far from Chicago. When we find the sensors that are close to Chicago, we name it as the `EPA.Points` sensors we need.

```{r filter}
test <- EPA.Points
test<- test[-c(24,25), ]
str(test@data)
plot(test)
EPA.Points <- test
```


## Generate the IDW

To set up the grid, we define the sample grid based on the extent of the `EPA.Points` file.

```{r}
grid <-spsample(EPA.Points, type = 'regular', n = 10000)
```

Then we run the IDW function from the `gstat` package for the variable of interest in `EPA.Points`.

```{r}
idw <- idw(EPA.Points$X2017MPC ~ 1, EPA.Points, newdata = grid)

idw.output = as.data.frame(idw)
names(idw.output)[1:3] <- c("long", "lat", "prediction")
```

Then we create the spatial points data frame.

```{r}
spg <- idw.output
coordinates(spg) <- ~ long + lat
```

The last step is to coerce the grid to Spatial Pixels Data Frame and to raster.

```{r}
gridded(spg) <- TRUE

raster_idw <- raster(spg)
```

## Set the Projection

We use the following code to set the standard of the projection.

```{r}
projection(raster_idw) <- CRS("+init=epsg:4326")
```

## Plot the IDW Results

Firstly, we mask raster by Chicago polygon file.

```{r}
masked_idw <- mask(raster_idw, Chicago)
```

Then we plot the masked raster.

```{r}
tm_shape(masked_idw) + tm_raster("prediction", style = "quantile", n = 100, legend.show = FALSE)
```

The previous raster file only shows the IDW result. Then We plot the masked raster with point emissions.

```{r}
tm_shape(masked_idw) + tm_raster("prediction", style = "quantile", n = 100, legend.show = FALSE) +
  tm_shape(EPA.Points) + tm_bubbles(size = "X2017MPC", col = "X2017MPC", palette = "Blues", 
                                    scale = 0.5, style = "quantile", legend.size.show = FALSE, title.col = "2017 PM2.5 Mean Concentration")

tmap_mode("view")

tm_shape(masked_idw) + 
  tm_raster("prediction", style = "quantile", 
            n = 100, alpha = 0.4, legend.show = FALSE) + 
  tm_shape(EPA.Points) + 
  tm_bubbles(col = "X2017MPC", 
             palette = "Blues", 
             scale = .5, 
             style = "quantile", 
             title.col = "2017 Mean PM2.5") 

tmap_mode("plot")
```

From the code above, we use `tm_bubbles` to plot the sensors. The function allows us to change colors, sizes, and display styles of the sensor. For example, we applied the color (`palette`) as `"Blues"` for our map, and we could change it to any colors from  `RColorBrewer::display.brewer.all`. Also, we use .5 for the `scale` and `"quantile"` as the `style`. We could increase the scale to enlarge the EPA sensors on the map and change the style as one of "fixed", "sd", "equal", "pretty", "quantile", "kmeans", "hclust", "bclust", or "fisher". If we change the color to `"Greens"` and the style to `"sd"`, we could get the following map:

```{r}
tm_shape(masked_idw) + 
  tm_raster("prediction", style = "quantile", 
            n = 100, alpha = 0.4, legend.show = FALSE) + 
  tm_shape(EPA.Points) + 
  tm_bubbles(col = "X2017MPC", 
             palette = "Greens", 
             scale = .6, 
             style = "sd", 
             title.col = "SD of 2017 Mean PM2.5") 
```

We can also scoll the map to zoom in and zoom out. Moreover, `tm_dots` is also a similar option for ploting some dot locations on the map, and it has almost the same coding structure as `tm_bubbles`. For more features of these commands, We can find the document of `tm_bubbles` from this link here: https://www.rdocumentation.org/packages/tmap/versions/1.2/topics/tm_bubbles.

We plot the masked raster with Community Area overlay added.

```{r}
tm_shape(masked_idw, bbox=Chicago) + 
  tm_raster("prediction", style = "quantile", pal = "Blues", n = 5, legend.show = TRUE, title = "PM2.5 prediction") + 
  tm_shape(Chicago) + tm_borders(alpha=.5,) + 
  tm_layout(frame = F, legend.position = c("left","bottom"))

```

## Combine the shapefiles and the PM2.5 values

Extracting the Pixel Values from the Raster for chicago Community Areas and removing NAs, we have the cleaned masked IDW results.

```{r}
chi.ave <- raster::extract(masked_idw, Chicago)

chi.ave <- lapply(chi.ave, na.omit)
```

Then we calculate the average value by Community Areas.

```{r}
ca.ave <- lapply(chi.ave, FUN=mean)
ca.ave  <- unlist(ca.ave )
```

We attach the Mean Values to Community Area information by grabing Community Area names/IDs from Chicago shapefiles. 

```{r}
chi.info <- cbind(as.integer(as.character(Chicago@data$area_numbe)))
head(chi.info)
```

Then we attach EPA Average numbers to the data frame.

```{r}
chica.final <- cbind(chi.info, ca.ave)
colnames(chica.final) <- c("area_numbe", "PM2.5")
as.data.frame(chica.final)

Chicago<- merge(Chicago, chica.final, by = "area_numbe")
head(Chicago@data)
```

Now we plot the masked raster with Community Area overlay.

```{r}
tm_shape(Chicago) + 
  tm_fill("PM2.5", style = "quantile", n = 5, pal = "Blues",legend.show = TRUE, title = "PM2.5 by CA") + 
  tm_shape(Chicago) + tm_borders(alpha=.5) + 
  tm_layout(frame = F, legend.position = c("left","bottom"))
```

## Save the Results

Finally, we write the data to a CSV file. 

```{r}
write.csv(chica.final, "PM5_aggregated.csv")
```

